Batch Bayesian Optimization via Multi-objective Acquisition Ensemble for Automated Analog Circuit Design

### 2018-02-08 paper/conclusion.tex
minor revision of experimental results and conclusion

In this paper, a batch Bayesian optimization algorithm is proposed for the
automation of analog circuit design. The parallelization is achieved via
the multi-objective ensemble of acquisition function. In each iteration, the
candidate points are sampled from the Pareto front of multiple acquisition
functions. We compared the proposed MACE algorithm using analytical benchmark
functions and real-world circuits, it is shown that the MACE algorithm is
competitive compared with state-of-the-art method listed in the paper.

>>>>

In this paper, a batch Bayesian optimization algorithm is proposed for the
automation of analog circuit design. The parallelization is achieved via
the multi-objective ensemble of acquisition functions. In each iteration, the
candidate points are sampled from the Pareto front of multiple acquisition
functions. We compared the proposed MACE algorithm using analytical benchmark
functions and real-world circuits, it is shown that the MACE algorithm is
competitive compared with the state-of-the-art methods listed in the paper.

### 2018-02-08 paper/experiments.tex
minor revision of experimental results and conclusion

\begin{table*}[t]
    \centering
    \caption{Optimization results of the benchmark functions}
    \label{tab:result_analytical}
    \begin{tabular}{lllllllllll}
        \toprule
                & Eggholder                &  Branin              &  Alpine1              & Hartmann6                   \\ \midrule
        MACE-1  & 87.65$\pm$75.83          &  1.05e-5$\pm$1.31e-5          & 2.66305$\pm$1.05844            & 0.0646869$\pm$0.0621189 \\
        LCB-1   & 153.9$\pm$112.8          &  6.86e-5$\pm$1.13e-4          & 5.66812$\pm$1.76973            & 0.125565$\pm$0.122684   \\
        EI-1    & 172.8$\pm$132.2          &  1.62e-2$\pm$1.63e-2          & 2.46061$\pm$1.56079            & 0.110561$\pm$0.146809   \\
        MACE-4  & 46.38$\pm$40.89          &  \textbf{4.62e-6$\pm$6.64e-6} & \textbf{0.903805$\pm$0.835209} & \textbf{0.0275738$\pm$0.052254}  \\
        BLCB-4  & 56.86$\pm$35.91          &  4.32e-5$\pm$6.33e-5          & 1.8843$\pm$0.938873            & 0.06447$\pm$0.0621176   \\
        EI-LP-4 & \textbf{44.68$\pm$56.45} &  2.11e-2$\pm$1.84e-2          & 1.0059$\pm$0.456865            & 0.0540446$\pm$0.0558557 \\
        qKG-4   & 106.4$\pm$67.64          &  2.65e-1$\pm$2.70e-1          & 3.01513$\pm$1.13414            & 0.47134$\pm$0.18939     \\
        qEI-4   & 72.13$\pm$52.08          &  3.29e-4$\pm$1.14e-3          & 2.7074$\pm$1.05145             & 0.186088$\pm$0.116323   \\
        \hline
                & Ackley2                       &  Rosenbrock2               & Ackley10              & Rosenbrock10        \\ \midrule
        MACE-1  & 1.71474$\pm$1.12154           &  0.026173$\pm$0.051189              & 3.1348$\pm$0.447874           & 499.697$\pm$300.899 \\
        LCB-1   & 1.624$\pm$0.926437            &  0.0201124$\pm$0.0205367            & 3.14797$\pm$0.519164          & 517.944$\pm$288.955 \\
        EI-1    & 1.0136$\pm$0.985858           &  13.5508$\pm$9.52734                & 18.8006$\pm$0.652136          & 1367.08$\pm$637.507 \\
        MACE-4  & 1.07906$\pm$0.886466          &  \textbf{0.00095416$\pm$0.00093729} & \textbf{2.56439$\pm$0.535488} & \textbf{158.116$\pm$50.0024} \\
        BLCB-4  & 1.40051$\pm$1.02849           &  0.00191986$\pm$0.00180895          & 3.27543$\pm$0.735501          & 406.819$\pm$127.351 \\
        EI-LP-4 & \textbf{0.284265$\pm$0.24634} &  2.73645$\pm$2.05923                & 18.2682$\pm$0.608564          & 721.351$\pm$327.365 \\
        qKG-4   & 5.59394$\pm$1.80595           &  5.03976$\pm$3.72014                & 18.197$\pm$0.764103           & 705.112$\pm$412.762 \\
        qEI-4   & 2.87373$\pm$1.02405           &  10.1881$\pm$15.0432                & 18.3686$\pm$0.501869          & 655.208$\pm$340.954 \\
        \bottomrule
    \end{tabular}
\end{table*}

The proposed MACE algorithm was tested using eight benchmark functions and two
real-world circuits. Four state-of-the-art parallel Bayesian optimization
methods were compared, including the BLCB algorithm, the local penalization
method with EI acquisition function, the qEI and qKG methods\footnote{We
implemented the BLCB algorithm as the availble open source implementations only
allow discrete input; for the EI-LP method, the code is downloaded from
https://github.com/SheffieldML/GPyOpt is used; the code for qEI and qKG is
downloaded from https://github.com/wujian16/Cornell-MOE}.

@@

The mean convergence plot is shown in Figure~\ref{fig:CovPlotBenchmark}. The
statistics of the final optimization results are listed in
Table~\ref{tab:result_analytical}. As can be seen, when running with batch size
$B = 4$, the MACE algorithm gives the best performance for six out of the eight
benchmarks.  For the 2D Ackley function, EI-LP is the best algorithm, while for
the Eggholder function, the MACE, BLCB, and EI-LP have quite similar
performance.  Compare the batch MACE and the sequential MACE, the speedup is
also dramatic.

% \begin{table*}[tp]
%     \tiny
%     \centering
%     \caption{Optimization results of the benchmark functions}
%     \label{tab:result_analytical}
%     \begin{tabular}{lllllllllll}
%         \toprule
%         Function    & MACE-1             & LCB-1                 &  EI-1                 &  MACE-4               & BLCB-4                &  EI-LP-4              & qKG-4                   & qEI-4                     \\ \midrule
%         Eggholder   & 88$\pm$76          &  1.5e+2$\pm$1.1e2     &  1.7e2$\pm$1.3e2      &  46$\pm$41            &  57$\pm$36            &  45$\pm$56            &  1.1e2$\pm$68  &  72$\pm$52            & \\
%         Branin      & 1e-5$\pm$1.3e-5    &  6.9e-5$\pm$1.1e-4    &  0.016$\pm$0.016      &  4.6e-6$\pm$6.6e-6    &  4.3e-5$\pm$6.3e-5    &  0.021$\pm$0.018      &  0.27$\pm$0.27 &  3.3e-4$\pm$1.1e-3    & \\
%         Alpine1     & 2.7$\pm$1.1        &  5.7$\pm$1.8          &  2.5$\pm$1.6          &  0.9$\pm$0.84         &  1.9$\pm$0.94         &  1$\pm$0.46           &  3$\pm$1.1     &  2.7$\pm$1.1          & \\
%         Hart6       & 0.065$\pm$0.062    &  0.13$\pm$0.12        &  0.11$\pm$0.15        &  0.028$\pm$0.052      &  0.064$\pm$0.062      &  0.054$\pm$0.056      &  0.47$\pm$0.19 &  0.19$\pm$0.12        & \\
%         Ackley      & 1.7$\pm$1.1        &  1.6$\pm$0.93         &  1$\pm$0.99           &  1.1$\pm$0.89         &  1.4$\pm$1            &  0.28$\pm$0.25        &  5.6$\pm$1.8   &  2.9$\pm$1            & \\
%         Rosenbrock  & 0.026$\pm$0.051    &  0.02$\pm$0.021       &  14$\pm$9.5           &  9.5e-4$\pm$9.4e-4    &  1.9e-3$\pm$1.8e-3    &  2.7$\pm$2.1          &  5$\pm$3.7     &  10$\pm$15            & \\
%         G05         & 3.1$\pm$0.45       &  3.1$\pm$0.52         &  19$\pm$0.65          &  2.6$\pm$0.54         &  3.3$\pm$0.74         &  18$\pm$0.61          &  18$\pm$0.76   &  18$\pm$0.5           & \\
%         G07         & 500$\pm$300        &  520$\pm$290          &  1400$\pm$640         &  160$\pm$50           &  410$\pm$130          &  720$\pm$330          &  710$\pm$410   &  660$\pm$340          & \\
%         \bottomrule
%     \end{tabular}
% \end{table*}

@@

The operational amplifier~\cite{wang2014enabling} shown in
Figure~\ref{fig:schDAC2014} is used for comparison, the circuit is designed
using the 180nm process. It has ten design parameters, including the lengths
and widths of transistors, the resistance of resistors and the capacitance of
capacitors. The circuit is simulated using the HSPICE circuit simulator.

We want to maximize the gain, unit gain frequency (UGF) and the phase margin (PM) for this amplifier, the $FOM$ is constructed as:
$$
\mathit{FOM} = -1.2 \times \mathit{gain} - 10 \times \mathit{UGF} - 1.6 \times \mathit{PM}
$$

For this circuit, we compared the MACE algorithm with the BLCB and EI-LP
algorithms, the qKG and qEI are not compared as the calculation of qEI and qKG
acquisition functions become very slow for the ten-dimensional functions.

@@

We run the algorithms in sequential mode and batch mode, for the batch
mode, batch size is set to $B = 4$. The number initial random sampling is set
to $N_{init} = 100$, and the number of iterations is set to $N_{iter} = 100$.

\begin{table}[h]
    \centering
    \caption{Optimization results of the operational amplifier}
    \label{tab:result_opamp}
    \begin{tabular}{llll}
        \toprule
        Algorithm & Results     \\ \midrule
        MACE-1    & 0.0$\pm$0.0 \\
        LCB-1     & 0.0$\pm$0.0 \\
        EI-1      & 0.0$\pm$0.0 \\
        MACE-4    & 0.0$\pm$0.0 \\
        BLCB-1    & 0.0$\pm$0.0 \\
        EI-LP-4   & 0.0$\pm$0.0 \\
        \bottomrule
    \end{tabular}
\end{table}

The mean convergence plot for the sequential and batch runs are plotted in
Figure~\ref{fig:resDAC2014}. As can be seen, the MACE algorithm gave better
solution compared with other algorithms, the sequential MACE algorithm found
better solution than BLCB and EI-LP with batch size $B = 4$, which showed that
the multi-objective acquisition ensemble is more robust than relying on single
acquisition function; compared with the sequential MACE algorithm, the MACE
algorithm with batch size $B = 4$ showed considerable speedup.

@@

\begin{figure}[htbp]
    \begin{center}
        \centerline{\includegraphics[width=\columnwidth]{./img/classE.eps}}
        \caption{Schematic of the power amplifier}
        \label{fig:schPA}
    \end{center}
\end{figure}

The power amplifier shown in Figure~\ref{fig:schPA} is used for comparison, the
circuit is designed using the 180nm process with twelve design parameters, the
circuit is simulated by HSPICE to get its performances.

For this power amplifier, we aim to maximize the power added efficiency (PAE) and the output power (Pout), the $FOM$ is constructed as
$$
\mathit{FOM} = -3 \times \mathit{PAE} - \mathit{Pout}
$$

@@

The MACE, BLCB, and EI-LP algorithms were tested in both sequential and batch
mode. The number of initial sampling is $N_{init} = 100$, the number of
iterations is $N_{iter} = 100$, the batch size is set to $B = 4$, so the total
number of HSPICE simulations is 500 for each batch run and 200 for each
sequential run.

\begin{figure}[htbp]
\vskip 0.2in
\begin{center}
\centerline{\includegraphics[width=\columnwidth]{./img/ClassE_mean.eps}}
\caption{Optimization results of the class-E power amplifier}
\label{fig:resClassE}
\end{center}
\vskip -0.2in
\end{figure}

The mean convergence plot is given in Figure~\ref{fig:resClassE}. We can see
that the MACE outperformed the BLCB and EI-LP in both sequential and batch
mode. For the batch runs, the MACE convergences fastest among the three
algorithms, while the sequential MACE has similar performance as the batch
EI-LP method.

>>>>

The proposed MACE algorithm was tested using eight benchmark functions and two
real-world circuits. Four state-of-the-art parallel Bayesian optimization
methods were compared, including the BLCB algorithm, the local penalization
method with EI acquisition function, the qEI and qKG methods\footnote{We
implemented the BLCB algorithm as the available open source implementations only
allow discrete input. For the EI-LP method, the code is downloaded from
https://github.com/SheffieldML/GPyOpt. The code for qEI and qKG is
downloaded from https://github.com/wujian16/Cornell-MOE}.

@@

The mean convergence plot is shown in Figure~\ref{fig:CovPlotBenchmark}. The
statistics of the final optimization results are listed in
Table~\ref{tab:result_analytical}. As can be seen, when running with batch size
$B = 4$, the MACE algorithm gives the best performance for six out of the eight
benchmarks.  For the 2D Ackley function, EI-LP is the best algorithm, while for
the Eggholder function, the MACE, BLCB, and EI-LP have quite similar
performance.  Compared with the sequential MACE, the speedup of the batch MACE is
also dramatic.

\begin{table*}[tp]
    \centering
    \caption{Optimization results of the benchmark functions}
    \label{tab:result_analytical}
    \begin{tabular}{lllllllllll}
        \toprule
        Function           & MACE-1 & LCB-1  & EI-1  & MACE-4  & BLCB-4  & EI-LP-4 & qKG-4 & qEI-4  \\ \midrule
         Branin            &        &        &       &         &         &         &       &        \\
         Alpine1           &        &        &       &         &         &         &       &        \\
         Hartmann6         &        &        &       &         &         &         &       &        \\
         Eggholder         &        &        &       &         &         &         &       &        \\
         Ackley2           &        &        &       &         &         &         &       &        \\
         Rosenbrock2       &        &        &       &         &         &         &       &        \\
         Ackley10          &        &        &       &         &         &         &       &        \\
         Rosenbrock10      &        &        &       &         &         &         &       &        \\
        \bottomrule
    \end{tabular}
\end{table*}

@@

\textcolor{red}{TODO: if possible, run qKG and qEI for this circuit}

The operational amplifier~\cite{wang2014enabling} shown in
Figure~\ref{fig:schDAC2014} is used to test the Bayesian optimization algorithms. The circuit is designed
using the 180nm process. It has 10 design parameters, including the lengths
and widths of transistors, the resistance of the resistors and the capacitance of the
capacitors. The circuit is simulated using the commercial HSPICE circuit simulator.

We want to maximize the gain, unit gain frequency (UGF) and the phase margin (PM) for this amplifier, the Figure of Merit $FOM$ is constructed as:
$$
\mathit{FOM} = -1.2 \times \mathit{gain} - 10 \times \mathit{UGF} - 1.6 \times \mathit{PM}
$$

For this circuit, we compared the MACE algorithm with the BLCB and EI-LP
algorithms. The qKG and qEI are not compared as the computation of qEI and qKG
acquisition functions become very slow for the ten-dimensional functions.

@@

We run the algorithms in sequential mode and batch mode. For the batch
mode, the batch size is set to $B = 4$. The number of initial random sampling is set
to $N_{init} = 100$, and the number of iterations is set to $N_{iter} = 100$.

\begin{table}[]
    \centering
    \caption{Optimization results of the operational amplifier}
    \label{tab:result_opamp}
    \begin{tabular}{llll}
        \toprule
        Algorithm & Results     \\ \midrule
        MACE-1    & 0.0$\pm$0.0 \\
        LCB-1     & 0.0$\pm$0.0 \\
        EI-1      & 0.0$\pm$0.0 \\
        MACE-4    & 0.0$\pm$0.0 \\
        BLCB-1    & 0.0$\pm$0.0 \\
        EI-LP-4   & 0.0$\pm$0.0 \\
        \bottomrule
    \end{tabular}
\end{table}

The mean convergence plot for the sequential and the batch runs are plotted in
Figure~\ref{fig:resDAC2014}. As can be seen, the MACE algorithm gave better
solution compared with other algorithms. The sequential MACE algorithm found
better solution than BLCB and EI-LP with batch size $B = 4$, which showed that
the multi-objective acquisition ensemble is more robust than relying on single
acquisition function. Compared with the sequential MACE algorithm, the MACE
algorithm with batch size $B = 4$ showed considerable speedup.

@@

\begin{figure}[htbp]
    \begin{center}
        \centerline{\includegraphics[width=\columnwidth]{./img/classE.eps}}
        \caption{Schematic of the power amplifier\textcolor{red}{TODO: replace this schematic with eps file} }
        \label{fig:schPA}
    \end{center}
\end{figure}

The power amplifier shown in Figure~\ref{fig:schPA} is used to test the Bayesian optimization algorithms. The
circuit is designed using the 180nm process with 12 design parameters, the
circuit is simulated by the commercial HSPICE circuit simulator to get its performances.

For this power amplifier, we aim to maximize the power added efficiency (PAE) and the output power (Pout), the Figure of Merit $FOM$ is constructed as
$$
\mathit{FOM} = -3 \times \mathit{PAE} - \mathit{Pout}
$$

@@

The MACE, BLCB, and EI-LP algorithms were tested in both sequential and batch
modes. The number of initial sampling is $N_{init} = 100$. The number of
iterations is $N_{iter} = 100$. The batch size is set to $B = 4$. The total
number of HSPICE simulations is 500 for each batch run and 200 for each
sequential run.

\begin{figure}[htbp]
\vskip 0.2in
\begin{center}
\centerline{\includegraphics[width=\columnwidth]{./img/ClassE_mean.eps}}
\caption{Optimization results of the class-E power amplifier.}
\label{fig:resClassE}
\end{center}
\vskip -0.2in
\end{figure}

The mean convergence plot is given in Figure~\ref{fig:resClassE}. We can see
that the MACE outperformed the BLCB and EI-LP in both sequential and batch
mode. For the batch runs, the MACE converges fastest among the three
algorithms, while the sequential MACE has similar performance as the batch
EI-LP method.
